How Slow do my Transient Type Checks Go?

### 2019-08-17 paper.tex
Revise The Benchmarks section. Cut it down to what’s needed for the paper
Signed-off-by: Stefan Marr <git@stefan-marr.de>

\begin{document}
\title{How Slow do my Transient Type checks Go?}

@@

\subsection{The Benchmarks}
In our previous work \cite{roberts-and-co-ecoop-2019}, we evaluated the performance of transient type checks using 21 benchmarks. We started with the 14 benchmarks of the \AWFY
benchmark suite \cite{Marr2016}: Bounce, CD, DeltaBlue, GraphSearch, Havlak, Json, List, Mandelbrot, NBody, Permute, Richards, Sieve, Storage and Towers benchmarks. We added another nine benchmarks taken from the gradual-typing literature.
Our goal was to complement the \AWFY with additional ones that are used for similar experiments and can be ported to Grace.
We surveyed a number of papers
\citep{Takikawa2016,Vitousek2017,Muehlboeck2017,Bauman2017,Richards2017,Stulova2016,Greenman2018}
and selected additional benchmarks that have been used by multiple papers.
Two of these benchmarks (NBody and Sieve) overlapped with the \AWFY suite,
or were available in different versions.
While not always behaviourally equivalent,
we chose the \AWFY versions since we already used them to
establish a performance baseline.
The selected benchmarks as well as the papers in which they were used are shown in
Table~\ref{tab:gradual-benchmarks}.

\begin{table}[htb]
  \caption{Benchmarks selected from literature.}
  \label{tab:gradual-benchmarks}
  \begin{center}
    \begin{tabular}{l l r}
      Fannkuch & \cite{Vitousek2017,Greenman2018} \\
      Float & \cite{Vitousek2017,Muehlboeck2017,Greenman2018} \\
      Go & \cite{Vitousek2017,Muehlboeck2017,Greenman2018} \\
      NBody & \cite{Kuhlenschmidt:2018:preprint,Vitousek2017,Greenman2018} & used \cite{Marr2016} \\
      Queens & \cite{Vitousek2017,Muehlboeck2017,Greenman2018} & used \cite{Marr2016} \\
      PyStone & \cite{Vitousek2017,Muehlboeck2017,Greenman2018} \\
      Sieve & \cite{Takikawa2016,Muehlboeck2017,Bauman2017,Richards2017,Greenman2019jfp} & used \cite{Marr2016} \\
      Snake & \cite{Takikawa2016,Muehlboeck2017,Bauman2017,Richards2017,Greenman2019jfp} \\
      SpectralNorm & \cite{Vitousek2017,Muehlboeck2017,Greenman2018} \\
    \end{tabular}
  \end{center}
\end{table}

The benchmarks were modified to have complete type information.  To
ensure correctness and completeness of these experiments, we ensured that each benchmark is completely typed.  To assess the performance
overhead of type checking, we compared the execution of Moth with all
checks disabled against an execution that has all checks enabled.  The
results of this evaluation demonstrated that Moth's transient type checking has an overhead of 5\% (min.\ -13\%,
max. 79\%) compared to Moth's peak performance with type checking disabled.

We did not measure programs that mixed typed and untyped code because
with our implementation technique we expected fully typed programs to
have the largest overhead, which is not the case for other kinds of
gradual type checking.

@@

In order to identify those type annotations, we generated coloured
scatter plots for each benchmark based on whether each individual type
annotation was present in that configuration or not. The position of
the dots in each coloured plot are the same as Figure~\ref{f:full}, but
they are coloured as to whether each annotation is there or not: This resulted in 1,775 scatterplots --- one for each type annotation for each benchmark. \kjx{if we say this, we should show a bunch for one benchmark.}
We then manually inspected all these scatterplots,
but unfortunately most of these showed no discernible pattern.
\kjx{YEAH BUT SOME DID. Shouldn't we show at least one?}

@@

The high-performance computing community has been investigating how
tools and visualization can help developers to utilize their systems
more efficiently \citep{Papenhausen:2016:IVT,daSilva:2019:PSV}.
Their focus is typically on parallelisation opportunities,
guided by runtime feedback, cost models, or heuristics. 
Their large body of work \citep{Isaacs:2014:PerfViz} uses various approaches,
though, we are not aware of work that has used an approach similar to ours.
One may argue, this is largely a question of practicality.\sm{
i put this in to hint at issues with usability.}
\kjx{Hmm not sure what the last sentence here means!}

@@

In the future, we hope to investigate statistical techniques to
determine the significance of each type annotation's contribution to a
programs overall performance. We would also like to investigate 
whether this approach can assist with optimsations for programmers' day-to-day  development, or help VM engineers identifying performance bugs in the underlying virtual machines.
\sm{is our approach practical. Is it useful for day-to-day development?
What's its context, use, general application?}
\kjx{filled out the last sentence}

>>>>

\begin{document}
\title{How Slow do my Transient Type Checks Go?}

@@

\subsection{The Benchmarks}

For this work, we rely on the benchmark suite compiled for previous work
\cite{roberts-and-co-ecoop-2019}.
It is a collection of 21 benchmarks in total,
derived from the \AWFY
benchmark suite \cite{Marr2016} and other benchmarks
from the gradual-typing literature.

% In our previous work , we evaluated the performance of transient type checks using 21 benchmarks. We started with the 14 benchmarks of the : Bounce, CD, DeltaBlue, GraphSearch, Havlak, Json, List, Mandelbrot, NBody, Permute, Richards, Sieve, Storage and Towers benchmarks. We added another \ugh{nine}\sm{14 plus 9 isn't 21} benchmarks taken from the gradual-typing literature.\sm{it's also unclear whether this is about previous work or what we do now.}
% Our goal was to complement the \AWFY with additional ones that are used for similar experiments and can be ported to Grace.
% We surveyed a number of papers
% \citep{Takikawa2016,Vitousek2017,Muehlboeck2017,Bauman2017,Richards2017,Stulova2016,Greenman2018}
% and selected additional benchmarks that have been used by multiple papers.
% Two of these benchmarks (NBody and Sieve) overlapped with the \AWFY suite,
% or were available in different versions.
% While not always behaviourally equivalent,
% we chose the \AWFY versions since we already used them to
% establish a performance baseline.
% The selected benchmarks as well as the papers in which they were used are shown in
% Table~\ref{tab:gradual-benchmarks}.
% \sm{this paragraph needs to be cleaned up.
% We need to say what we use now. Clearly, and not intermingled with previous work.
% And then, afterwards we can say whether this is the same we did before. or how it differs.
% }
%
% \begin{table}[htb]
%   \caption{Benchmarks selected from literature.}
%   \label{tab:gradual-benchmarks}
%   \begin{center}
%     \begin{tabular}{l l r}
%       Fannkuch & \cite{Vitousek2017,Greenman2018} \\
%       Float & \cite{Vitousek2017,Muehlboeck2017,Greenman2018} \\
%       Go & \cite{Vitousek2017,Muehlboeck2017,Greenman2018} \\
%       NBody & \cite{Kuhlenschmidt:2018:preprint,Vitousek2017,Greenman2018} & used \cite{Marr2016} \\
%       Queens & \cite{Vitousek2017,Muehlboeck2017,Greenman2018} & used \cite{Marr2016} \\
%       PyStone & \cite{Vitousek2017,Muehlboeck2017,Greenman2018} \\
%       Sieve & \cite{Takikawa2016,Muehlboeck2017,Bauman2017,Richards2017,Greenman2019jfp} & used \cite{Marr2016} \\
%       Snake & \cite{Takikawa2016,Muehlboeck2017,Bauman2017,Richards2017,Greenman2019jfp} \\
%       SpectralNorm & \cite{Vitousek2017,Muehlboeck2017,Greenman2018} \\
%     \end{tabular}
%   \end{center}
% \end{table}

Each benchmark has a complete set of type annotations,
\ie, all possible type annotations are present,
which is verified in the parser.

In our previous work, we \cite{roberts-and-co-ecoop-2019}
we determined that the overhead of type checking on Moth is on average
of 5\% (min.\ -13\%, max. 79\%).
This compares the peak performance of Moth with all
checks disabled against an execution that has all checks enabled. 
%  The
% results of this evaluation demonstrated that Moth's transient type checking has an overhead  compared to Moth's peak performance with type checking disabled.
%
% We did not measure programs that mixed typed and untyped code because
% with our implementation technique we expected fully typed programs to
% have the largest overhead, which is not the case for other kinds of
% gradual type checking.

@@

In order to identify those type annotations, we generated coloured
scatter plots for each benchmark based on whether each individual type
annotation was present in that configuration or not. The position of
the dots in each coloured plot are the same as Figure~\ref{f:full}, but
they are coloured as to whether each annotation is there or not: This resulted in 1,775 scatterplots --- one for each type annotation for each benchmark. We then manually inspected all these scatterplots, though almost all of these graphs showed no discernible pattern, we were able to identify some for about a third of the benchmarks.

@@

The high-performance computing community has been investigating how
tools and visualization can help developers to utilize their systems
more efficiently \citep{Papenhausen:2016:IVT,daSilva:2019:PSV}.
Their focus is typically on parallelisation opportunities,
guided by runtime feedback, cost models, or heuristics. 
Their large body of work \citep{Isaacs:2014:PerfViz} uses various approaches,
though, we are not aware of work that has used an approach similar to ours.
One may argue, this is largely a question of practicality.

@@

In the future, we hope to investigate statistical techniques to
determine the significance of each type annotation's contribution to a
programs overall performance. We would also like to investigate 
whether this approach can assist with optimsations for programmers' day-to-day  development, or help VM engineers identifying performance bugs in the underlying virtual machines.
